---
title: "renderAmpCurves and renderMeltCurves"
author: "Konstantin A. Blagodatskikh"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{pcrPlateInput overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Introduction

**renderAmpCurves** and **renderMeltCurves** function represent amplification and melting data from real-time PCR experiments as curves based on **plotly** package.
Main advantage of using this functions instead of regular plot renders is that it glued with **RDML** package. Minimal usage recures only `GetFData(long.table = TRUE)` function output. Also it have interactive feature - fast curves hiding without total plot redraw.

```{r, results='hide', message=FALSE, warning=FALSE}
library(shinyMolBio)
library(tidyverse)
library(RDML)
# load RDML file
rdml <- RDML$new(system.file("/extdata/stepone_std.rdml", package = "RDML"))
```
```{r, results='asis'}
# create Shiny UI
renderAmpCurves(inputId = "firstLook", # Shiny input ID
              label = "Example", # optional plot label 
              ampCurves = rdml$GetFData(long.table = TRUE), # Amplification curves
              interactive = FALSE
) 
```

# Curves customization

## Color

Curve color can be directly provided by adding column **color** to **ampCurves** table or by choosing column that defines color with **colorBy** param.

```{r, results='asis'}
# create Shiny UI
renderAmpCurves(inputId = "color1",
              ampCurves = rdml$GetFData(long.table = TRUE),
              colorBy = "sample", # sample name will define color
              interactive = FALSE
) 

renderAmpCurves(inputId = "color2",
              ampCurves = rdml$GetFData(long.table = TRUE) %>% 
                mutate(color = "red"), # All curves will be red
              interactive = FALSE
)
```

## Linetype

Curve linetype can be setted by choosing column that defines linetype with **linetypeBy** param.

```{r, results='asis'}
# create Shiny UI
renderAmpCurves(inputId = "linetype",
              ampCurves = rdml$GetFData(long.table = TRUE),
              linetypeBy = "sample.type", # sample.type will define color
              interactive = FALSE
) 
```

# Show Markers

You can show *Cq* or *Tm* values on curves as markers setting `showCq = TRUE` or `showTm = TRUE`. Then input table have to contain *cq* or *tm* column.

```{r, results='asis'}
# create Shiny UI
renderAmpCurves(inputId = "cq", 
              ampCurves = rdml$GetFData(
                rdml$AsTable(cq = data$cq), # Get Cq values from file
                long.table = TRUE), 
              showCq = TRUE, # Add Cq markers to curves
              colorBy = "sample",
              interactive = FALSE
) 
```

# Hiding curves

Individual curves can be hidden without plot redraw. Use *updateCurves* function with *fdata.name* as *hideCurves* param.
Run `shinyMolBio::runExample("pcrPlateInput")` to see this in action.