---
title: "renderAmpCurves and renderMeltCurves"
author: "Konstantin A. Blagodatskikh"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{renderAmpCurves and renderMeltCurves overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Introduction

**renderAmpCurves** and **renderMeltCurves** function represent amplification and melting data from real-time PCR experiments as curves based on **plotly** package.
Main advantage of using this functions instead of regular plot renders is that it glued with **RDML** package. Minimal usage recures only `GetFData(long.table = TRUE)` function output. Also it have interactive feature - fast curves hiding without total plot redraw.

```{r, results='hide', message=FALSE, warning=FALSE}
library(shinyMolBio)
library(tidyverse)
library(RDML)
library(chipPCR)
# load RDML file
rdml <- RDML$new(system.file("/extdata/stepone_std.rdml", package = "RDML"))
```
```{r, results='asis'}
renderAmpCurves(inputId = "firstLook", # Shiny input ID
              label = "First Look", # optional plot label 
              ampCurves = rdml$GetFData(long.table = TRUE), # Amplification curves
              interactive = FALSE
) 
```

# Curves customization

## Color

Curve color can be directly provided by adding column **color** to **ampCurves** table or by choosing column that defines color with **colorBy** param.

```{r, results='asis'}
renderAmpCurves(inputId = "color1",
                "Color by Sample Name",
              ampCurves = rdml$GetFData(long.table = TRUE),
              colorBy = "sample", # sample name will define color
              interactive = FALSE
) 

renderAmpCurves(inputId = "color2",
                "All 'red'",
              ampCurves = rdml$GetFData(long.table = TRUE) %>% 
                mutate(color = "red"), # All curves will be red
              interactive = FALSE
)
```

## Linetype

Curve linetype can be setted by choosing column that defines linetype with **linetypeBy** param.

```{r, results='asis'}
renderAmpCurves(inputId = "linetype",
                "Different Linetypes",
              ampCurves = rdml$GetFData(long.table = TRUE),
              linetypeBy = "sample.type", # sample.type will define color
              interactive = FALSE
) 
```

# Show Markers

You can show *Cq* or *Tm* values on curves as markers setting `showCq = TRUE` or `showTm = TRUE`. Then input table have to contain *cq* or *tm* column.

```{r, results='asis'}
renderAmpCurves(inputId = "cq",
                "Show Cq Values",
              ampCurves = rdml$GetFData(
                rdml$AsTable(cq = data$cq), # Get Cq values from file
                long.table = TRUE), 
              showCq = TRUE, # Add Cq markers to curves
              colorBy = "sample",
              interactive = FALSE
) 
```

# Show Cq Threshold Line

Threshold lines can be shown by choosing column that splits different threshold values with **thBy** param. Then input table have to contain *quantFluor* column.

```{r, results='asis'}
# Create function for curves preprocessing
dataType$set("public", "Process",
             function(thValue) {
               # Subtract background
               private$.adp$fpoints$fluor <- 
                 CPP(self$adp$fpoints$cyc,
                     self$adp$fpoints$fluor,
                     bg.range = c(10,20))$y.norm
               # Calc Cq by threshold method
               self$cq <- th.cyc(self$adp$fpoints$cyc, self$adp$fpoints$fluor, r = thValue)[1, 1]
               # Write threshold value
               self$quantFluor <- thValue
             },
             overwrite = TRUE)

rdml <- RDML$new(system.file("/extdata/lc96_bACTXY.rdml", package = "RDML"))

# Manual threshold values for different targets
thValues <- c("bACT" = 0.03, "X" = 0.05, "Y" = 0.04, "IPC" = 0.01)

# Preprocess every curve
for (react in rdml$experiment[[1]]$run[[1]]$react) {
  for (fdata in react$data) {
    fdata$Process(thValues[fdata$tar$id])
  }
}


tbl <- rdml$AsTable(quantFluor = data$quantFluor, # Add threshold values to table
                    cq = data$cq)
renderAmpCurves("th", 
                "Show Thershold Lines", 
                rdml$GetFData(tbl, long.table = TRUE), 
                colorBy = "target",
                thBy = "target", # Add threshold lines (separated by targets)
                interactive = FALSE) 
```

# Custom Plotly Code

You can add custom qouted plotly code by **plotlyCode** parameter. Note that you
have to add value **p** inside your quoted code to link it with render output.

```{r, results='asis'}
markTbl <- tbl %>% 
  filter(position %in% c("D03", "D07"),
         target == "bACT")

renderAmpCurves("th", 
                "Show Thershold Lines", 
                rdml$GetFData(tbl, long.table = TRUE), 
                colorBy = "target",
                plotlyCode = quote(
                  # Add Cq values for tubes D03 and D07 for target bACT as blue points
                  add_markers(p,
                              data = markTbl,
                              name = ~sample,
                              x = ~cq,
                              y = ~quantFluor,
                              marker = list(color = "blue",
                                            size = 15)) %>%
                    # Set background color to light yellow
                    layout(paper_bgcolor = '#ffffe0',
                           plot_bgcolor = '#ffffe0')
                ),
                interactive = FALSE)
```

# Melting curves

**renderMeltCurves** function provides all functionality described previosly in **renderAmpCurves** examples. Differences are *showTm* param instead of *showCq* and there is no *thBy* param.

```{r, results='asis'}
# load RDML file
rdml <- RDML$new(system.file("/extdata/BioRad_qPCR_melt.rdml",
                             package = "RDML"))
mdps <- rdml$GetFData(dp.type = "mdp", long.table = TRUE)
mdps[, diffFluor := c(0, diff(fluor)) * -1, by = fdata.name]

renderMeltCurves("melt", 
                "Show melting curves",
                mdps,
                fluorColumn = "diffFluor",
                colorBy = "target",
                interactive = FALSE)
```

# Hiding and highlighting curves

Individual curves can be hidden without plot redraw. Use *updateCurves* function with *fdata.name* as *hideCurves* param. Or highlighted with *fdata.name* as *highlightCurves* param.
Run `shinyMolBio::runExample("pcrPlateInput")` to see this in action. 
Curves hiding occures after wells selection at PCR plate and higlighting after 
mouse hovering above PCR plate or details table.
