---
title: "renderADplot"
author: "Konstantin A. Blagodatskikh"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{renderADplot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Introduction

**renderADplot** function visualizes standard allelic discrimination plot based on **plotly** package.
Main advantage of using this functions instead of regular plot renders is that it glued with **RDML** package. Minimal usage recures only `GetFData(long.table = TRUE)` function output. Also it has interactive features.

Standard allelic discrimination plot is a plot with two axis: one is signal for allele 1 and second is signal for allele 2. Signal can be end point RFU, Cq or another countable parametr. Column name that contains axis splitting is determined by **targetColumn** argument. Column name that contains signal level is determined by **valueColumn** argument. If **targetColumn** contains more than two unique values you can specify which to use by **xAxisTarget** and **yAxisTarget**.

```{r, results='hide', message=FALSE, warning=FALSE}
library(shinyMolBio)
library(tidyverse)
library(RDML)
# load RDML file
rdml <- RDML$new(system.file("/extdata/test.rdml", package = "shinyMolBio"))
tbl <- rdml$AsTable(endPointRFU = mean(tail(data$adp$fpoints$fluor, 5)))
tbl <- group_by(tbl, position)
# generate genotypes by simple rule based on end point RFU signal:
# Reaction1 > 400 - AA
# Reaction2 > 400 - GG
# Reaction1 and Reaction2 > 400 - AG
tbl <- mutate(tbl,
              genotype = paste(
                if (endPointRFU[1] > 400 && endPointRFU[2] > 400) "AG"
                else if (endPointRFU[1] > 400) "AA"
                else if (endPointRFU[2] > 400) "GG"
                else "NA"
              ))
```
```{r, results='asis'}
renderADplot(inputId = "firstLook", 
             label = "AD plot with default arguments", 
             adData = tbl,
             targetColumn = "target.dyeId",
             xAxisTarget = "FAM",
             yAxisTarget = "HEX", 
             valueColumn = "endPointRFU",
             colorBy = "genotype",
             interactive = FALSE)
```

# Curves customization

## Color

Curve color can be directly provided by adding column **color** to **adData** table or by choosing column that defines color with **colorBy** param.

```{r, results='asis'}
renderADplot(inputId = "color1", 
             label = "Color by Sample Name", 
             adData = tbl,
             colorBy = "sample",
             interactive = FALSE)

renderADplot(inputId = "color1", 
             label = "Color by Sample Name", 
             adData = tbl %>% 
               mutate(color = "red"),
             interactive = FALSE)
```

## Polar coordinates

Curve linetype can be setted by choosing column that defines linetype with **linetypeBy** param.

You can view plot in polar coordinates with angle as X axis and distance as Y axis by 
setting **polar = TRUE**.

```{r, results='asis'}
renderADplot(inputId = "polar", 
             label = "Polar Coordinates", 
             adData = tbl,
             polar = TRUE,
             interactive = FALSE)
```


# Custom Plotly Code

You can add custom qouted plotly code by **plotlyCode** parameter. Note that you
have to add value **p** inside your quoted code to link it with render output.

```{r, results='asis'}
centers <- tbl %>% 
  group_by(target.dyeId, genotype) %>% 
  summarise(endPointRFU = mean(endPointRFU)) %>% 
  spread(target.dyeId, endPointRFU) %>% 
  filter(genotype != "NA")

renderADplot("th", 
                "Show Thershold Lines", 
                tbl,
                plotlyCode = quote(
                  # add markers to the centers of genotypes
                  add_markers(p,
                              data = centers,
                              name = ~genotype,
                              x = ~FAM,
                              y = ~HEX,
                              marker = list(color = "blue",
                                            size = 50,
                                            opacity = 0.2)) %>%
                    # Set background color to light yellow
                    layout(paper_bgcolor = '#ffffe0',
                           plot_bgcolor = '#ffffe0')
                ),
                interactive = FALSE)
```



# Hiding and highlighting curves

Individual points can be hidden without plot redraw. Use *updateADplot* function with *fdata.name* as *hidePoints* param. Or highlighted with *fdata.name* as *highlightPoints* param.
Run `shinyMolBio::runExample("pcrPlateInput")` to see this in action. 
Points hiding occures after wells selection at PCR plate and higlighting after 
mouse hovering above PCR plate or details table.
